The GWR wildfire smoke's association with 2013 Oregon medication of short acting beta agonists

### 2017-09-01 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

---
title: "The GWR Smoke's effect on Oregon 2013 Medication of Beta-2 agonists"
author: "Jingyang Liu"
date: "May 17, 2017"
output: html_document
---

```{r library, include=FALSE, echo=FALSE}
library(tidyverse)
# library(data.table)
library(survival)
library(ggplot2)
library(htmlTable)
library(knitr)
library(broom)
library(lme4)
library(maps)
library(lubridate) # date
library(ggrepel)
library(sp)

@@

## Overview (1, 2)

1. Intro: We found an association with wildfire smoke and respiratory outcomes in Washington state 2012 using a novel estimate of smoke concentration, geo-graphically weighted ridge regression (GWR). We want to know if smoke impacts other health outcomes, specifically respiratory rescue medications (beta 2 agonists) (abbreviate to SABA)

2. Research question: We evaluated the association between smoke concentrations using the GWR method and SABA refills in Oregon state during the 2013 wildfire season.

@@

## Method Description (3)

3.Methods: use case-crossover for all health outcome figures and results. Make sure smoke is assigned via zipcode. For the descriptive of smoke, I think the county-level for now. We may move to zip.

@@

```{r DLM for disease, }

@@

start <- Sys.time()
for(j in var_list) { # begin first loop of variable names (outcomes)

  j <- "broken_arm"
  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

@@

which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

@@

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_geo_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_geo_lag_B + strata(personkey), 
              or_disease_lag_df)

@@

@@

}

@@

```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 25

>>>>

---
title: "The GWR wildfire smoke's association with 2013 Oregon medication of short acting beta agonists"
author: "Jingyang Liu"
date: "May 17, 2017"
output: html_document
---

```{r library, include=FALSE, echo=FALSE}
library(tidyverse)
# library(data.table)
library(survival)
library(ggplot2)
library(htmlTable)
library(knitr)
library(broom)
library(lme4)
library(maps)
library(lubridate) # date
library(ggrepel)
library(sp)
library(cowplot)

@@

## Overview

**1. Introduction: **
We found an association with wildfire smoke and respiratory outcomes in Washington state 2012 using a novel estimate of smoke concentration, geo-graphically weighted ridge regression (GWR). We want to know if smoke impacts other health outcomes, specifically respiratory rescue medications (beta 2 agonists) (abbreviate to SABA)

**2. Research question: **
We evaluated the association between smoke concentrations using the GWR method and SABA refills in Oregon state during the 2013 wildfire season.

@@

**3.Methods: **
use case-crossover for all health outcome figures and results. Make sure smoke is assigned via zipcode. For the descriptive of smoke, I think the county-level for now. We may move to zip.

@@

```{r DLM for disease respiratory, echo = FALSE, message=FALSE, warning=FALSE}

@@

  j <- c('respiratory')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

@@

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

@@

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

@@

DLestimate$lag <- 0:lag_max
p1 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("Respiratory") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

@@

```{r DLM for disease asthma, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('asthma')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

# DLestimate
DLestimate$lag <- 0:lag_max

p2 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("Asthma") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease pneumonia, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('pneumonia')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max
p3 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("Pneumonia") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease acute_bronch, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('acute_bronch')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max

p4 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("Acute bronch") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease copd, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('copd')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max
p5 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("COPD") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease cvd, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('cvd')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max
p6 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("cvd") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease isch_heart_dis, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('isch_heart_dis')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max
p7 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("Isch heart disease") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease arrhythmia, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('arrhythmia')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max
p8 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("Odds Ratio for 10ug/m^3 Increase in PM2.5") +
  ggtitle("Arrhythmia") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease heart_failure, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('heart_failure')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max
p9 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("Heart failure") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease cerbrovas_dis, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('cerbrovas_dis')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max
p10 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("Cerbrovas disease") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease myocardial_infarc, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('myocardial_infarc')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max
p11 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("Myocardial infarc") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r DLM for disease broken_arm, echo = FALSE, message=FALSE, warning=FALSE}

  j <- c('broken_arm')

  # standard for statement without parallel computing
  #for(j in 1:length(var_list)){
  read_path <- paste('../data_new/casecrossover/time_strat/or', j, 'may_to_sep_time_strat_casecross.csv', sep='_')
  or_disease <- read_csv(read_path)

# which(colnames(or_disease_lag_df)=="geo_smk_pm10_zip") #174
# which(colnames(or_disease_lag_df)=="geo_smk_pm10_lag1_zip") #175
# which(colnames(or_disease_lag_df)=="wrf_temp_zip") #94
# which(colnames(or_disease_lag_df)=="wrf_temp_lag1_zip") #151

or_disease_geo_lag_B <- or_disease_geo_lag%*%B
or_disease_temp_lag_B <- or_disease_temp_lag%*%B

fit <- clogit(outcome ~ or_disease_geo_lag_B + or_disease_temp_lag_B + strata(personkey), 
              or_disease_lag_df)

DLestimate$lag <- 0:lag_max
p12 <- ggplot(DLestimate, aes(x=lag, y=estimate, ymin=lower, ymax=upper)) + 
  geom_ribbon(alpha=.5) + geom_line(size=2) + 
  geom_line(aes(x=lag), linetype=2, size=2) + 
  theme_bw() + 
  xlab("lagged days of events") + 
  ylab("OR for 10ug/m^3 Increase") +
  ggtitle("Broden arm") +
  geom_hline(yintercept=1, linetype="dashed", color = "red")

```{r}
plot_grid(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, ncol=4)

@@

```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, echo=FALSE}
or_ndc_county_ts_df <- or_ndc_county_ts %>%
  select(1:4, pop, rate_per_100k, day, geo_smk_pm_county,
         geo_smk_pm_lag1_county:geo_smk_pm_lag7_county, wrf_temp_county,
         wrf_temp_lag1_county:wrf_temp_lag7_county) %>%
  mutate(geo_smk_pm10_county = geo_smk_pm_county/10,
         geo_smk_pm10_lag1_county =geo_smk_pm_lag1_county/10,
         geo_smk_pm10_lag2_county =geo_smk_pm_lag2_county/10,
         geo_smk_pm10_lag3_county =geo_smk_pm_lag3_county/10,
         geo_smk_pm10_lag4_county =geo_smk_pm_lag4_county/10,
         geo_smk_pm10_lag5_county =geo_smk_pm_lag5_county/10,
         geo_smk_pm10_lag6_county =geo_smk_pm_lag6_county/10,
         geo_smk_pm10_lag7_county =geo_smk_pm_lag7_county/10)

which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 25
